How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2022-09-07 Draft_September_2022/Full_draft_20220901.rmd
#corrections on references #problem construction -> problem-construction #added acknowledgements and declaration of interest #knitting nicely!

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "1 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
    ```
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about, and act towards, the environment transnationally. Still, these remain understudied in this context. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency of the Amazon as a topic in presidential speeches inside and outside Brazil. While constructing the Amazon a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest presidents are from the Amazon, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

# Prepare data for problem constructions

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest on the ground [@campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different sites, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (UNFCCC Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move away from the Amazon region. That is, presidents, talk to the economic and social needs of the people domestically while boasting about environmental policy internationally.

This article proceeds as follows: first, we propose a conceptual framework to understand problem-construction in discourses. We then review Amazonian historiography literature to identify the main phases and their underlying problem construction. In the methodology section, we explain how the framework is operationalized and present the codebook. We proceed to describe, visualize, and interpret our findings in the analysis section. Finally, we conclude with a discussion connecting our findings to past and present environmental policies.

@@

```{r Figure 2, fig.cap="Amazonian speeches in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.18, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.34, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "August Fires", size=3) +
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year (n=6130)",
       subtitle = "We annotate five local maxima with highest frequencies.",
       color='Y series') +
  theme_am()
Amazon_year
```

@@

```{r Figure 3, fig.cap="Deforestation rates since 1988", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
def_rates <- ggplot(amazon_def_year, aes(x = year)) +
  geom_line(aes(y = area), color= "darkolivegreen", linetype="dotted", size= 1.5)+
  geom_smooth(se=FALSE, color="black", aes(y=area), linetype="solid", size= 0.75)+
  annotate("text", x = 2004, y = 29000,
           label = "2004 Plan to Combatand Control Deforestation", size = 3) +
  annotate("text", x = 2012.5, y = 3500, label = "2012 Forest Code ", size = 3) +
  labs(x = "", y = "Deforestation in squared kilometers",
       title = "Deforestation rates since 1988",
       subtitle= "Data from INPE, Brazilian Ministry of Science and Technology.",
       color = "",
       caption= "Black curve smoothed using loess.")+
  theme_am()
def_rates
```

@@

### Pure-type problem constructions

We conceptualize four problem constructions: sovereignty, economic integration, social development, and conservation. Figure 4, below, illustrates the share of each pure-type in time. Pure problem constructions dominate, with an average of 56% of all Amazonian statements across time.We identify a discursive shift in problem constructions from developing the Amazon economically, to conserving it and providing services to its citizens. Pure economic integration constructions dominated constructions from 1985 to the late 2000s. This is especially pertinent during the Cardoso administration. Brazilian presidents after Cardoso moved from employing one problem-constructions consistently to a more diverse strategy. During Lula’s terms in office, constructing the Amazon purely as an issue of economic integration decreased, while constructing the Amazon as issues of environmental conservation and social development increased. Conservation and social development constructions surpassed economic integration around 2010 and continued to dominate as Amazonian constructions during the Rousseff administration. Still, environmental conservation and social development problem-constructions were not as dominant in terms of frequency or time in discourses as economic integration was until the early 2000s. Finally, we observe a steady increase in sovereignty constructions in the late 2000s, reaching its highest frequencies with the presidencies of Bolsonaro.

```{r Figure 4, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size =1, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  theme(legend.position = "bottom")
pure_types_time
```

@@

### Mixed-type problem constructions

At the level of the speech presidents might mix multiple problem constructions. Although presidents prefer pure problem-constructions, they also construct the Amazon as a multifaceted issue. Mixed-type problem constructions in discourse offer more intricate understandings of the Amazon as a problem. Constructing the Amazon as multiple issues averages at 18% of all constructions in time. Some mixed-types rarely appear so we focus our discussion on those mixed-types with higher incidence. Figure 5, below, displays the averages of mixed-types constructions by presidents.

@@

The most frequent mixed-type problem construction of the Amazon is economic conservation. This mixed-type construction, composed of Amazonian statements that construct the Amazon as a problem of both economic integration and environmental conservation, generally increased in time. Its increase over time suggests that later presidents, along with being more diverse in how they constructed the amazon as an issue (see Figure 4), increasingly find ways to reconcile the dictum between the economy and the environment that prevailed in the previous years. The second most common mixed-type mixes social development and economic integration and appears, on average, in 5.4% of all statements. This is not surprising, as developing countries like Brazil repeatedly claim a “right to develop” when it comes to negotiating strong climate commitments and policies with a focus on both social development and economic integration. Nonetheless, there is a general increase of mixed-types in time, which is expected given the rise of global agendas understanding interconnections of social, environmental, and economic domains such as the Millennium Development Goals and the Sustainable Development Goals.

@@

```{r Figure 6, fig.cap="Percent of problem constructions across locations", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure Sovereignty", 1, 0))
# model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:24)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  geom_vline(xintercept = c(1933.78, 2771.49, 5075.48, 8306.76), linetype = "dotted") +
  ggplot2::annotate(geom = "text", x = c(1933.78,2771.49 , 5075.48, 8306.76),
                    y = c(0.45,0.5,0.47,0.5),
                    label = c("Brasilia", "Sao Paulo", "New York", "Paris"), alpha = 0.4) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted share of each problem-construction as a function of distance from the Amazon",
       subtitle="Sovereignty was removed because of statistical insignificance (table 1).", 
       caption = "Curve in the plot is estimated using loess method.") +
    theme_am() +
  theme(legend.position = "bottom")
all_loc
```

@@

```{r Table 1, fig.cap="Logistic Regressions Output", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
# summary statistics
md <-stargazer::stargazer(model_logit_con, model_logit_ei, model_logit_sd,
                          model_logit_sov, digits = 10, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon (km)",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE)
```

@@

This paper investigates how the Amazon has been constructed as a problem in Brazilian presidential speeches since 1985. Our findings are threefold: we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (UNFCCC Summits). Second, the same government adopts a diverse range of problem-construction. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, temporarily surpassing economic integration as a more common problem construction from 2010 to 2015. This implies an overall movement from only exploiting the Amazon for economic purposes to exploiting it and protecting it.  Constructing the Amazon as an issue of sovereignty became more pertinent after 2010. Finally, we find that presidents are more likely to construct the Amazon as an environmental conservation problem than economic integration or social development, as presidents move away from the Amazon region. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these relate to policies and environmental outcomes. Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as an issue in presidential discourse.

@@

\newpage
# References

<div id="refs"></div>

@@

codebook <- tibble::tibble("Problem Construction" = c("National Sovereignty", "Economic Integration","Social Development", "Environmental Conservation"),

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "7 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
    ```
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about, and act towards, the environment transnationally. Still, these remain understudied in this context. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency of the Amazon as a topic in presidential speeches inside and outside Brazil. While constructing the Amazon a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest presidents are from the Amazon, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

@@

# Prepare data for problem-constructions

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest on the ground [@campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different sites, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (UNFCCC Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move away from the Amazon region. That is, presidents, talk to the economic and social needs of the people domestically while boasting about environmental policy internationally.

This article proceeds as follows: first, we propose a conceptual framework to understand problem-construction in discourses. We then review Amazonian historiography literature to identify the main phases and their underlying problem-construction. In the methodology section, we explain how the framework is operationalized and present the codebook. We proceed to describe, visualize, and interpret our findings in the analysis section. Finally, we conclude with a discussion connecting our findings to past and present environmental policies.

@@

```{r Figure 2, fig.cap="Amazonian speeches in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.18, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.34, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992.5, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "2019 Record August Fires", size=3) +
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year",
       subtitle = "Sample composed by all 6130 speeches since 1985",
       color='Y series') +
  theme_am()
Amazon_year
```

@@

```{r Figure 3, fig.cap="Deforestation rates since 1988", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
def_rates <- ggplot(amazon_def_year, aes(x = year)) +
  geom_line(aes(y = area), color= "darkolivegreen", linetype="twodash", size= 0.5)+
  geom_smooth(se=FALSE, color="black", aes(y=area), linetype="solid", size= 0.5)+
  annotate("text", x = 2005, y = 29000,
           label = "2004 Plan to Combat and Control Deforestation", size = 3) +
  annotate("text", x = 2012.5, y = 3500, label = "2012 Forest Code ", size = 3) +
  labs(x = "", y = "Deforestation in squared kilometers",
       title = "Deforestation rates since 1988",
       subtitle= "Data from INPE, Brazilian Ministry of Science and Technology.",
       color = "",
       caption= "Black curve smoothed using loess.")+
  theme_am()
def_rates
```

@@

### Pure-type problem-constructions

We conceptualize four problem-constructions: sovereignty, economic integration, social development, and conservation. Figure 4, below, illustrates the share of each pure-type in time. Pure problem-constructions dominate, with an average of 56% of all Amazonian statements across time.We identify a discursive shift in problem-constructions from developing the Amazon economically, to conserving it and providing services to its citizens. Pure economic integration constructions dominated constructions from 1985 to the late 2000s. This is especially pertinent during the Cardoso administration. Brazilian presidents after Cardoso moved from employing one problem-constructions consistently to a more diverse strategy. During Lula’s terms in office, constructing the Amazon purely as an issue of economic integration decreased, while constructing the Amazon as issues of environmental conservation and social development increased. Conservation and social development constructions surpassed economic integration around 2010 and continued to dominate as Amazonian constructions during the Rousseff administration. Still, environmental conservation and social development problem-constructions were not as dominant in terms of frequency or time in discourses as economic integration was until the early 2000s. Finally, we observe a steady increase in sovereignty constructions in the late 2000s, reaching its highest frequencies with the presidencies of Bolsonaro.

```{r Figure 4, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>%
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am() +
  theme(legend.position = "bottom")
pure_types_time
```

@@

### Mixed-type problem-constructions

At the level of the speech presidents might mix multiple problem-constructions. Although presidents prefer pure problem-constructions, they also construct the Amazon as a multifaceted issue. Mixed-type problem-constructions in discourse offer more intricate understandings of the Amazon as a problem. Constructing the Amazon as multiple issues averages at 18% of all constructions in time. Some mixed-types rarely appear so we focus our discussion on those mixed-types with higher incidence. Figure 5, below, displays the averages of mixed-types constructions by presidents.

@@

The most frequent mixed-type problem-construction of the Amazon is economic conservation. This mixed-type construction, composed of Amazonian statements that construct the Amazon as a problem of both economic integration and environmental conservation, generally increased in time. Its increase over time suggests that later presidents, along with being more diverse in how they constructed the amazon as an issue (see Figure 4), increasingly find ways to reconcile the dictum between the economy and the environment that prevailed in the previous years. The second most common mixed-type mixes social development and economic integration and appears, on average, in 5.4% of all statements. This is not surprising, as developing countries like Brazil repeatedly claim a “right to develop” when it comes to negotiating strong climate commitments and policies with a focus on both social development and economic integration. Nonetheless, there is a general increase of mixed-types in time, which is expected given the rise of global agendas understanding interconnections of social, environmental, and economic domains such as the Millennium Development Goals and the Sustainable Development Goals.

@@

```{r Figure 6, fig.cap="Logistic Regression predicted values", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure Sovereignty", 1, 0))
# model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:24)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted share of each problem-construction as a function of\ndistance from the Amazon",
       subtitle="Sovereignty was removed because of statistical insignificance (table 1).", 
       caption = "Curves in the plots are estimated using loess method.") +
    theme_am() +
  theme(legend.position = "bottom")+
  facet_wrap(~p_c, ncol=3)

all_loc
```

@@

```{r Table 1, fig.cap="Logistic Regressions Output", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
# summary statistics
md <-stargazer::stargazer(model_logit_con, model_logit_ei, model_logit_sd,
                          model_logit_sov, digits = 10, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon (km)",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE,
                          title = "Logistic Regressions Output")
```

@@

This paper investigates how the Amazon has been constructed as a problem in Brazilian presidential speeches since 1985. Our findings are threefold: we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (UNFCCC Summits). Second, the same government adopts a diverse range of problem-construction. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, temporarily surpassing economic integration as a more common problem-construction from 2010 to 2015. This implies an overall movement from only exploiting the Amazon for economic purposes to exploiting it and protecting it.  Constructing the Amazon as an issue of sovereignty became more pertinent after 2010. Finally, we find that presidents are more likely to construct the Amazon as an environmental conservation problem than economic integration or social development, as presidents move away from the Amazon region. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these relate to policies and environmental outcomes. Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as an issue in presidential discourse.

@@

# Acknowledgements

The authors are extremely grateful to Graziella Moraes Silva and James Hollway for their support. The authors would also like to thank Matias, Federico, Rodrigo, and Anna for valuable comments on earlier drafts of this paper.

# Disclosure Statement

The authors report there are no competing interests to declare.

# References

<div id="refs"></div>
\newpage

@@

codebook <- tibble::tibble("problem-construction" = c("National Sovereignty", "Economic Integration","Social Development", "Environmental Conservation"),
